---
title:
date: "January 15, 2018"
output:
  rmdformats::html_clean:
    highlight: kate
---

```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

# Syllabus (FIN 411) Financial Problems
<br>
<br>
<center>
<b>SUNY POLYTECHNIC INSTITUTE</b>

<b>SCHOOL OF BUSINESS ADMINISTRATION</b>
</center>
<br>
<br>
**Instructor**: Matthew Brigida, Ph.D.

**Office**: Donovan 1277

**Office Hours**: Tuesday/Thursday/Friday  10:30--11:30

**Email**:  matthew.brigida@sunyit.edu

**Class Location**:  Donovan 1157

**Class Day/Time**: TR, 0900-1030am

**Text**: [Case Studies in Finance 6th Edition, by Robert Bruner, Kenneth Eades, Michael Schill. ISBN-13: 978-0073382456 ISBN-10: 0073382450](https://www.amazon.com/dp/0073382450/ref=cm_sw_r_em_apa_i_Tm1fBbR1CT3EV)

**Supplementary Material**: [5MinuteFinance.org](https://www.5minutefinance.org/)

## Course Structure and Interaction

The goal of the course is to provide good solutions to the text cases.  Completing the cases requires a good knowledge of concepts/terms/measures learned in earlier courses.  To help you remember this material, I may reference materials I provide via the following media:  

-  Via BlackBaord Collaborate.  You can view them live (I'll let you know when I am recording), or under the `Recordings` section of Collaborate.

-  Short videos many of which are/will be posted on my YouTube channel: https://www.youtube.com/channel/UCwekb0vAK-FKaqPF5gfd0eQ/featured


## Description

This is a cases course for upper-level finance students. The learning outcomes for this course are summarized below:

#. Students will use various skills to evaluate complex financial problems, and ultimately provide a recommendation on a course of action.  Among these skills are the ability to analyze financial statements, perform time-value-of-money calculations, perform capital budgeting tasks, compute costs of capital, and manipulate corporate capital structure, among others.

#.  Students will learn to communicate complex topics in a clear and concise manner.

#.  Students will learn to calculate and present various measures of financial performance appropriate to specific types of firms.  

#.  In short, in this course you will start to complete analyses like you are an analyst.  Specifically, we are transitioning away from simple calculations, and toward holistic analyses which incorporate many calculations, and your knowledge of financial topics and the interdependence between variables.  

## Academic Honesty Policy

Academic dishonesty will not be tolerated in this class. Cheating
on quizzes, examinations, and other forms of dishonesty (e.g., plagiarism, collusion, and
falsification of data) will be dealt with in a serious and formal manner. The penalty for academic
dishonesty in this class will be course failure. That is, any student who is found to be cheating
or engaged in other academically dishonest behavior will be failed for this course for this
semester. Course withdrawals to avoid such a failure will not be permitted. As a student, you
have a responsibility to become familiar with the Academic Honesty Policy found in the Student
Rights, Regulations, and Procedures Handbook.

<!---
## Office Hours

I will be logged on the course throughout the week to answer questions and review the discussions.  If a few students have questions on the material, and there is a good time to meet, I'll log into BB Collaborate and record the answers.

Students do not need to access the course simultaneously, and therefore have a measure of flexibility regarding when they access the course.  However, students should access the course regularly to keep up with assignments, and to maintain a presence in the Discussion Room.  Students should also often check their SUNY Poly email account.
-->

<!---
## Lectures

In addition to the lectures I posted to YouTube, I will also record lectures via BB Collaborate.  I'll try and let you know when I am recording each lecture, so you can log in an interact.  However all that is required is that you watch the lecture at some point---meaning you are free to just watch the recording.

**You must complete and submit your computation questions in a spreadsheet.  If you submit your computation answers in Word, I will reduce your score.  You should also not treat your spreadsheet as a Word or text file.  That is answers should be written as formulas, and not as text strings.**

**If your spreadsheet is disorganized you will lose points. The computations and answers should be easily found and understood by anyone opening your spreadsheet. See the 'Note on Spreadsheet Design' at the end of this syllabus.**

## Brokerage Accounts

We will use paper trading accounts provided by Interactive Brokers. These
trading accounts are the exact same as the actual brokerage accounts---except the money isn’t real.
You will have access to (and real data from) stock, bond, commodity, and foreign exchange markets.
While the data alone is worth a fair amount of money, Interactive Brokers is offering the accounts to us
for free. Keep this in mind while following the directions to set up your account customer service will
be nonexistent. **If you lose your password, or forget your username, etc, you will not have an
account for the semester.** In this case you will have to trade in another student’s account, or mine.
Please pay attention while setting up your account, and write everything down.
-->

## Cases

Throughout the semester we'll work on a series of projects that will help us understand both financial theories, and also, *importantly*, practical considerations we are faced with when applying these theories in the real-world. Some notes on the projects:

-  Each case will require computations either in a spreadsheet, or programming language.  In this course you should probably use a spreadsheet (Excel or similar).  If you write it in a programming language (such as R or Python) you'll have to share the code so I can reproduce your results.

-  If you have questions on a project you are encouraged to post your spreadsheet/code to the relevant discussion board so I and other students can help.  You learn a great deal by helping others, and by seeing how others approach the problem.  Also don't worry too much about others copying you.  It is fairly easy to tell when one spreadsheet has been copied from another.  Moreover, since you posted yours, it would be clear that you were the one copied off of.

## Participation

Your participation grade is a function of both your attendance, and your participation in the discussion while you are attending.  If you miss a few classes (particularly for a good reason) your grade won't be adversely affected, however if you *consistently* do not attend I will lower your participation score.  Similarly, if you *consistently* do not participate in the class discussion I will lower your score.  

<!---
## DISCUSSION:

For each two-week period I will post a set of *End-of-Chapter Questions* from the text.  You must choose 1 of the questions from the set and post your answer to the discussion board.  Try and choose a question that has not yet been answered, but if all the questions have been answered then try and add unique information in your answer.

The subject line of your discussion post should be the in the format [chapter].[question number].  For example, if you are answering question 5 from chapter 3, your subject line should be '5.3'.  Also, be sure to have one post for each answer. Don't answer multiple questions in one post.

If the questions requires computation, then use Excel (or other spreadsheet) and post your spreadsheet as your answer.  

<!-- In addition to your answer, you must post 1 response to other students' answers.  So each week you will post 1 answer, and 1 response to other answers, for a total of 2 posts to the discussion board each period. -->

<!-- To receive credit your posts must be constructive, preferably somewhat insightful, and show you have done the reading.  If your post does not meet the preceding criteria, please don't bother posting it because it will just clutter the discussion board.   -->

<!-- A particular period's discussion will conclude Sunday at 9pm. -->

<!-- **2.  Q/A:** You should post finance related short questions and answers similar to what you find at https://github.com/finance-AI/data/blob/master/studentQA/studentQA_fall_2016.txt.  These are submission from an earlier class.  You should submit your Q/A in the same format. -->

<!--   This assignment will help you learn our Corporate Finance material by writing concise answers to finance questions.  In addition, your submitted Q/A will be used to train a finance Artificial Intelligence.  I will feed the Q/A into a deep neural network to train a chatbot that can answer finance questions---think Siri for finance.    -->

<!-- I have created a file for Q/A for this course https://github.com/finance-AI/data/blob/master/studentQA/studentQA_570_summer_2017.txt.  Submit your Q/A there.  Adding to the file is easy---[see this video for a short tutorial](https://www.youtube.com/watch?v=iVC9UKkaiko), and [This is an example of a link to a commit.](https://github.com/FinancialMarkets/5MinuteFinance/commit/8f12f63b546a80fdb04e787514d967f1a1c0725f) -->

<!-- If your GitHub username is similar to your real name, then you don't have to tell me you have submitted the Q/A (I'll see it).  If your username is different, then send me an email from your clarion email address with the Subject "FIN 570 Q/A" letting me know your real name and GitHub username.  Note, GitHub shows how many lines you have added, so there is no need to number your questions.  Also, don't label the questions and answers.  -->
-->

## Course Communication

All communication will be through Blackboard and email. If you have question you are strongly encouraged to post it to a discussion forum instead of emailing me.  This way, other students can get the benefit of the question and answer.  It also saves me from answering the same question many times via email, and frees me up to answer more questions and generally provide more effective instruction for you.

## Grading

<center>

```{r grading, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| Item | Points |
| ---- | ------ |
| Case Submissions | 70 |
| Participation (including Blackboard Discussion) | 30 |
| Total Points | 100 |
"
cat(tabl)
```

</center>

<center>
**Final grades will be assigned according to the following scale**:
</center>

- 90 - 100 A
- 80 - 89.9 B
- 70 - 79.9 C
- 60 - 69.9 D
- $<$ 60 F

> +/- grades may be assigned at the instructors discretion.

## General Notes

- All times referred to in this course are Eastern Standard---unless otherwise indicated.
- There will be no make up exams or extra points assignments.
- Cheating will result in prosecution to the fullest extent possible under university rules.
- You are responsible for material covered in the online discussion, as well as text material.
-  **Internet Access:** This course requires that you have regular access to the internet to submit work.  You should not take this course if you plan on being in an area with insufficient internet access. "My internet was down for a week" is not an acceptable reason to hand in late work.
-  **Adding or Dropping the Course:** To add or drop the course the student should consult the university guidelines and withdrawal dates. The course instructor is not involved in a student's adding or withdrawing from the course.
- **Software:** You will need word processing and spreadsheet software to take
this course. Common examples of such software are Microsoft Word and
Excel. However, there is no need to buy this software if you don't already
have it. There are many free (open-source) alternatives which are just as
good (and which allow you to save/read files as .doc(x), .pdf, and .xls(x)).
Some widely used free office suites are LibreOffice (http://www.libreoffice.org)
and OpenOffice (http://www.openoffice.org). Feel free to download and use
these. *In this course all word processed submissions should be in .pdf, and
all spreadsheets should be submitted as .xlsx.*


## Tentative Outline

Week 1: August 28th and August 30th
Week 2: September 4th and September 6th
Week 3: September 11th and September 13th
Week 4: September 18th and September 20th
Week 5: September 25th and September 27th
Week 6: October 2nd and October 4th
            October 6- Tuesday October 9th Mid-semester break
Week 7: October 11th (Thursday)
Week 8: October 16th and October 18th
Week 9: October 23rd and October 25th
Week 10: October 30th and November 1st
Week 11: November 6th and November 8th
Week 12: November 13th and November 15th
Week 13: November 20th
            Thanksgiving November 22nd -25th
Week 14: November 27th and November 29th
Week 15: December 4th and December 6th
Week 16: Finals Week

Weeks 1 & 2 Setting themes
Cases in module: Ben & Jerry

Weeks 3 & 4
FedEx vs. UPS

Week 5 & 6
Financial Analysis and Forecasting
Case: Value-Line Publishing

Weeks 7 & 8
Estimating the cost of capital
Case: Nike, Inc.

Weeks 9 & 10
Capital Budgeting and Resource Allocation
Case: Aurora Textile Corp. pg. 283


> Based on student input, we may add topics in portfolio theory such as those found in chapters 5 through 10, and chapter 24.

## Some Notes on Spreadsheet Design

You should construct your spreadsheet as if you were an analyst at a company, and you were going to submit the spreadsheet to upper management.  Therefore, getting the correct answer can be considered the minimal amount of work.  The spreadsheet should be easily readable and organized.  There are a couple of reasons why this is important: (1) management often will check some numbers (or maybe change a few inputs if they have more up to date information) and it will reflect very poorly on you if they have to search around through a muddled and ill-conceived spreadsheet; and (2) anyone should be able to pick up your spreadsheet and complete it if you are not there (vacation, sick, or hopefully promoted).  Following are a couple tips on spreadsheet design, though it is far from exhaustive.

- Hard-code as little as possible.  You want a few cells for your inputs, or a place where you put your data, and then every other cell is linked and feeds off of these input cells. This way, to update your spreadsheet you simply change the inputs or drop in new data.  
- Take the time to label cells, and put in appropriate comments if necessary  - though comments should not be used excessively. Also, it is common to change the cell color depending on whether it is hard-coded (an input) or a formula.  This way you (or anyone else) can immediately look at a cell and tell whether it is one in which you can type (an input).  Don't forget to include a key.
- It is often better to add tabs to a spreadsheet than continue calculations on one tab.  You can easily page through spreadsheet tabs with `Ctrl+Shift' and `Page-up' or `Page-down'.
- Pivot tables.  While we probably won't need them in this course, you should nonetheless get to know them.  Pivot tables are incredibly useful for summarizing data, and it is very possible you will be asked in an interview whether you are familiar with them.  Similarly, get to know VLOOKUP.
-  If you are inputting a long formula, then break the calculation into multiple cells.  This makes it much easier to tell where a mistake was made - and everyone always spends a fair amount of time looking for errors.
- Excel has many built in formulas which can be useful, however it is important that you understand what the formula is doing to use them.  Blindly applying a formula can lead to trouble.  For example, if you use the IRR() function on cash flows with multiple roots, the formula will return the first root it finds without signaling to you that there are other roots.  Also, there are Excel formulas that are flat out incorrect - in particular the NPV() function.  So, use a function if it saves time, but first be sure you know what the function is doing and verify it works.  That said, in my experience it is better (and faster) to input your own formula instead of using Excel's.  You often have to break the calculation into a couple of steps, but this can be done quickly, and the result is a spreadsheet that you know works and is easily auditable.
