---
title:
date:
output: 
  rmdformats::html_clean:
    highlight: kate
#output: word_document
always_allow_html: true
---

```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

# Syllabus (FIN 411) Financial Problems
<br>
<br>
<center>
<b>SUNY POLYTECHNIC INSTITUTE</b>

<b>SCHOOL OF BUSINESS ADMINISTRATION</b>
</center>
<br>
<br>
**Instructor**: Matthew Brigida, Ph.D.

**Office**: On Campus Office is Donovan 1277
<!-- Online Summer 2022 | -->

**Office Hours**: Online by Request:  See the *Office Hours* section below. <!-- Tuesday 4--6pm, and on-request I will create a BB Collaborate session where you can log in and ask questions. -->

**Email**:  matthew.brigida@sunypoly.edu

**Class Location**:  Online <!-- Donovan 1242 -->

**Class Day/Time**: Asynchronous <!-- Tuesday 6--9:30pm -->
<!-- Asynchronous Online. Recordings will be posted to Blackboard Collaborate. -->

**Text**: [Case Studies in Finance 6th Edition, by Robert Bruner, Kenneth Eades, Michael Schill. ISBN-13: 978-0073382456 ISBN-10: 0073382450](https://www.amazon.com/dp/0073382450/ref=cm_sw_r_em_apa_i_Tm1fBbR1CT3EV)

You can also use the 8th edition if you have it under two conditions.  First, you have to post to the discussion board and ask what cases in the 8th you can substitute.  Second, you understand there are presently no videos covering the 8th edition cases.

**Supplementary Material**: 

- [Financial-Education Site](https://financial-education.github.io)

- [Corporate Finance Data Sets](http://pages.stern.nyu.edu/~adamodar/New_Home_Page/datacurrent.html)  These data sets contain very useful information for case analyses.

## Course Structure and Interaction

The goal of the course is to provide good solutions to the text cases.  Completing the cases requires a good knowledge of concepts/terms/measures learned in earlier courses.  To help you remember this material, I may reference materials I provide via the following media:  

-  Recorded videos and in Brightspace.

-  Short videos many of which are/will be posted on my YouTube channel: https://www.youtube.com/channel/UCwekb0vAK-FKaqPF5gfd0eQ/featured

You should use the discussion boards in Brightspace for any questions on material or the videos.  

I will post office hours on request. Usually I will try and find a time that works for the most number of students.  However, since this an asynchronous course with no set meeting time, my times for office hours may not match your schedule.  In this case we will not meet.  Note, also I do not schedule one-on-one meetings---there is not enough time in the week.  The bottom line is that you signed up for an asynchronous course, the **definition of which is we do not meet**.  If you think you *must* meet to complete this course, you should take it in a synchronous modality. 

Also, prior to meeting in office hours be sure to review all course materials on the topic, and then first ask questions on the discussion boards.  If we can't answer the question on the discussion board then it makes sense to meet in office hours. However, if you show up to office hours and say something like, "I don't get it" I'll simply refer you to the course materials.  If you have actually made an effort to understand the material, then you should be able to ask me a clear and pointed question. 

## Description

This is a cases course for upper-level finance students. The learning outcomes for this course are summarized below:

#. Students will use various skills to evaluate complex financial problems, and ultimately provide a recommendation on a course of action.  Among these skills are the ability to analyze financial statements, perform time-value-of-money calculations, perform capital budgeting tasks, compute costs of capital, and manipulate corporate capital structure, among others.

#.  Students will learn to communicate complex topics in a clear and concise manner.

#.  Students will learn to calculate and present various measures of financial performance appropriate to specific types of firms.

#.  In short, in this course you will start to complete analyses like you are an analyst.  Specifically, we are transitioning away from simple calculations, and toward holistic analyses which incorporate many calculations, and your knowledge of financial topics and the interdependence between variables.  

## Academic Honesty Policy

Academic dishonesty will not be tolerated in this class. Cheating
on quizzes, examinations, and other forms of dishonesty (e.g., plagiarism, collusion, and
falsification of data) will be dealt with in a serious and formal manner. The penalty for academic
dishonesty in this class will be course failure. That is, any student who is found to be cheating
or engaged in other academically dishonest behavior will be failed for this course for this
semester. Course withdrawals to avoid such a failure will not be permitted. As a student, you
have a responsibility to become familiar with the Academic Honesty Policy found in the Student
Rights, Regulations, and Procedures Handbook.


## Office Hours

The first method of answering questions is via the discussion boards. Any questions that cannot adequately be answered on the discussion boards can be solved with office hours.  Throughout the work week I will be logged on the course to answer questions and review the discussions.  I generally do not answer questions over the weekend. Note, I will not help if you wait until the last minute to complete your work.  If you post questions on a case on the day it is due I will likely not answer. 

When corresponding about your question on the discussion boards, we may set virtual office hours to more adequately answer your question. However there is no guarantee given your schedule that we will be able to find a compatible time.  So please consider:

**This is an asynchronous course, which means we do not meet at the same time.  If you need to meet to complete this course you should take it in an in-person or synchronous modality.**

## Cases

Throughout the semester we'll work on a series of projects that will help us understand both financial theories, and also, *importantly*, practical considerations we are faced with when applying these theories in the real-world. Some notes on the projects:

-  Most case will require computations either in a spreadsheet, or programming language.  In this course you should probably use a spreadsheet (Excel or similar).  If you write it in a programming language (such as R or Python) you'll have to share the code so I can reproduce your results.

-  If you have questions on a project you are encouraged to post your spreadsheet/code to the relevant discussion board so I and other students can help.  You learn a great deal by helping others, and by seeing how others approach the problem.  Also don't worry too much about others copying you.  It is fairly easy to tell when one spreadsheet has been copied from another.  Moreover, since you posted yours, it would be clear that you were the one copied off of.

-  All case submissions you make should be your work and your work alone.  There is no group work in this course.  If two students hand in the same spreadsheet they will both receive a 0 for the case.

### Case Report Structure

For each case you will hand in a report with the following general headings:

**I. Executive Summary**

This section should provide background and context to the case.  For example, in the Ben and Jerry's case it is important to know the company is incorporated in Vermont, and the effect on goal of the firm.  Whatever context you provide, however, should be relevant to the problem and recommendation.  Just random facts wastes time.

**II. Problem Statement**

This is a concise clear statement of the problem.  It should be a sentence or two.  

**III. Recommendation and Analysis**

This section should **first** answer the question posed in the problem statement.  The answer should be the first line of this section.  Don't hide your recommendations within the analysis.

Following your recommendations should be supporting analyses.  These are often calculations, for example you may say Ben and Jerry's should accept a given offer because it provides the highest price per share. Note, if you simple do random calculations that don't support your recommendation, this it worse than providing no analysis---it wastes time.
	
**IV. Alternate Solutions**

Here you can offer other solutions, along with accompanying benefits or drawbacks.  This section is optional and should not be long.


**V. Supplementary Material (Excel spreadsheets, etc)**

Many of the cases will require supplementary spreadsheet calculations. Do not simply hand in a spreadsheet with your calculations with no explanation. Think about how you would submit an analysis to your boss while working as an analyst. You must explain why you did what you did.

Your spreadsheet should be easily readable and usable by anyone.  Spreadsheets don't just calculate answers, they present them.  Lastly, you must complete and submit your computation questions in a spreadsheet.  If you submit your computation answers in Word, I will reduce your score.


#### Other General Notes

Many of the cases will require supplementary spreadsheet calculations.  Do not simply hand in a spreadsheet with your calculations with no explanation.  Think about how you would submit an analysis to your boss while working as an analyst.  You must explain why you did what you did.

Also, often students will ask for the correct answer or way of doing something.  This is a cases (real-world problem) course, and as such there is not one unique right answer.  The important thing is not your answer, but rather that you clearly state your reasoning for taking a particular approach.

**All cases should be handed in by uploading them to the Brightspace Dropbox.  Do not send them via email**

<!-- ## Participation -->

<!-- Your participation grade is a function of both your attendance, and your participation in the discussion while you are attending.  If you miss a few classes (particularly for a good reason) your grade won't be adversely affected, however if you *consistently* do not attend I will lower your participation score.  Similarly, if you *consistently* do not participate in the class discussion I will lower your score. -->

## Grading

Your final grade is an average of your grades for each case.  All cases are worth an equal amount.

<center>
```{r grading, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| Item | Points |
| ---- | ------ |
| Case Submissions | 90 |
| Participation | 10 |
| Total Points | 100 |
"
cat(tabl)
```
</center>


<center>
**Final grades will be assigned according to the following scale**:
</center>

- 90 - 100 A
- 80 - 89.9 B
- 70 - 79.9 C
- 60 - 69.9 D
- $<$ 60 F

> +/- grades may be assigned at the instructors discretion.

## General Notes

- All times referred to in this course are Eastern Standard---unless otherwise indicated.
- There will be no make up exams or extra points assignments.
- Cheating will result in prosecution to the fullest extent possible under university rules.
- You are responsible for material covered in the online discussion, as well as text material.
-  **Internet Access:** This course requires that you have regular access to the internet to submit work.  You should not take this course if you plan on being in an area with insufficient internet access. "My internet was down for a week" is not an acceptable reason to hand in late work.
-  **Adding or Dropping the Course:** To add or drop the course the student should consult the university guidelines and withdrawal dates. The course instructor is not involved in a student's adding or withdrawing from the course.
- **Software:** You will need word processing and spreadsheet software to take
this course. Common examples of such software are Microsoft Word and
Excel. However, there is no need to buy this software if you don't already
have it. There are many free (open-source) alternatives which are just as
good (and which allow you to save/read files as .doc(x), .pdf, and .xls(x)).
Some widely used free office suites are LibreOffice (http://www.libreoffice.org)
and OpenOffice (http://www.openoffice.org). Feel free to download and use
these. *In this course all word processed submissions should be in .pdf, and
all spreadsheets should be submitted as .xlsx.*


## Tentative Outline

The case due dates are <b style="color:red">bold and highlighted in red</b> in the table below.


<!--- Regular semester schedule----
| Week | Case |
| ---- | ---- |
| Weeks 1 and 2: 8/30--<b style="color:red">9/10</b> | Setting themes--Cases in module: **Ben & Jerry** |
| Weeks 3 and 4: 9/13--<b style="color:red">9/24</b> | **FedEx vs. UPS** |
| Weeks 5 and 6: 9/27--<b style="color:red">10/8</b> | Krispy Kreme  |
| Weeks 7 and 8: 10/11--<b style="color:red">10/22</b> | Teletech | ## The Body Shop |
| Weeks 9 and 10: 10/25--<b style="color:red">11/5</b> | [Securitization](https://matt-brigida.github.io/securitization-presentation-abridged/) |
| Weeks 11 and 12: 11/8--<b style="color:red">11/19</b> | Value-Line Publishing (HD vs LOW) | 
| Weeks 13 and 14: 11/22--<b style="color:red">12/10</b> | Capital Budgeting and Resource Allocation--Case: **Aurora Textile Corp.** |
-->

<!--- Fall semester schedule---->
| Due | Case |
| ---- | ---- |
| <b style="color:red">September 12</b>  | Setting themes--Cases in module: Ben & Jerry |
| <b style="color:red">September 26</b>  | FedEx Corp vs. United Parcel Service |
| <b style="color:red">October 10</b>  | Krispy Kreme  (8th ed: Financial Detective 2016) |
| <b style="color:red">October 24</b> | Teletech (8th ed: H.J. Heinz) | ## The Body Shop |
| <b style="color:red">November 7</b> | [Securitization](https://matt-brigida.github.io/securitization-presentation-abridged/) |
| <b style="color:red">November 21</b> | Value-Line Publishing (HD vs LOW) (8th ed: Whole Foods)|
| <b style="color:red">December 10</b> | Capital Budgeting and Resource Allocation--Case: Aurora Textile Corp. |


<!--- Summer 2024 semester schedule---->
<!-- | Due | Case | -->
<!-- | ---- | ---- | -->
<!-- | <b style="color:red">June 8</b>  | Setting themes--Cases in module: **Ben & Jerry** | -->
<!-- | <b style="color:red">June 15</b>  | **FedEx vs. UPS** | -->
<!-- | <b style="color:red">June 22</b>  | Krispy Kreme  | -->
<!-- | <b style="color:red">June 29</b> | Teletech | ## The Body Shop | -->
<!-- | <b style="color:red">July 6</b> | [Securitization](https://matt-brigida.github.io/securitization-presentation-abridged/) | -->
<!-- | <b style="color:red">July 13</b> | Value-Line Publishing (HD vs LOW) |  -->

<!--- | <b style="color:red">July 19</b> | Capital Budgeting and Resource Allocation--Case: **Aurora Textile Corp.** or **TBD**| -->

<!-- Financial Analysis and Forecasting--Case: **Value-Line Publishing (HD vs LOW)** | -->
<!--- Estimating the cost of capital--Case: **Nike, Inc.** -->

- [Some notes on the cases are here.](https://github.com/SUNY-Poly-Finance/FIN_411_text_case_notes)
- [Krispy Kreme](https://suny_poly_finance.gitlab.io/fin_411_krispy_kreme_notes/)
- [Securitization](https://matt-brigida.github.io/securitization-presentation-abridged/)

We may substitute one of the above cases for the following:

-  [Teletech](https://suny_poly_finance.gitlab.io/fin_411_teletech_notes/)
-  [Deposit Insurance](https://financial-education.github.io/deposit_insurance_valuation_by_option/)
-  Machine Learning/Artificial Intelligence in Finance

## Policy on Artifical Intellegence and LLM Use

You may use LLMs to come up with ideas and check solutions. You may not use LLMs to write your case report or construct your spreadsheet.  Your case submission must be your own words.  If it is clear that LLMs have written part of your case, or completed a portion of your spreadsheet, you will receive a 0.

## Some Notes on Spreadsheet Design

You should construct your spreadsheet as if you were an analyst at a company, and you were going to submit the spreadsheet to upper management.  Therefore, getting the correct answer can be considered the minimal amount of work.  The spreadsheet should be easily readable and organized.  There are a couple of reasons why this is important: (1) management often will check some numbers (or maybe change a few inputs if they have more up to date information) and it will reflect very poorly on you if they have to search around through a muddled and ill-conceived spreadsheet; and (2) anyone should be able to pick up your spreadsheet and complete it if you are not there (vacation, sick, or hopefully promoted).  Following are a couple tips on spreadsheet design, though it is far from exhaustive.

- Hard-code as little as possible.  You want a few cells for your inputs, or a place where you put your data, and then every other cell is linked and feeds off of these input cells. This way, to update your spreadsheet you simply change the inputs or drop in new data.  
- Take the time to label cells, and put in appropriate comments if necessary  - though comments should not be used excessively. Also, it is common to change the cell color depending on whether it is hard-coded (an input) or a formula.  This way you (or anyone else) can immediately look at a cell and tell whether it is one in which you can type (an input).  Don't forget to include a key.
- It is often better to add tabs to a spreadsheet than continue calculations on one tab.  You can easily page through spreadsheet tabs with `Ctrl+Shift' and `Page-up' or `Page-down'.
- Pivot tables.  While we probably won't need them in this course, you should nonetheless get to know them.  Pivot tables are incredibly useful for summarizing data, and it is very possible you will be asked in an interview whether you are familiar with them.  Similarly, get to know VLOOKUP.
-  If you are inputting a long formula, then break the calculation into multiple cells.  This makes it much easier to tell where a mistake was made - and everyone always spends a fair amount of time looking for errors.
- Excel has many built in formulas which can be useful, however it is important that you understand what the formula is doing to use them.  Blindly applying a formula can lead to trouble.  For example, if you use the IRR() function on cash flows with multiple roots, the formula will return the first root it finds without signaling to you that there are other roots.  Also, there are Excel formulas that are flat out incorrect - in particular the NPV() function.  So, use a function if it saves time, but first be sure you know what the function is doing and verify it works.  That said, in my experience it is better (and faster) to input your own formula instead of using Excel's.  You often have to break the calculation into a couple of steps, but this can be done quickly, and the result is a spreadsheet that you know works and is easily auditable.

## Some Discussion Papers on AI

You are not required to read these.

+ [McKinsey AI](https://www.mckinsey.com/featured-insights/artificial-intelligence)
+ [AI: The Next Digital Frontier?](https://www.mckinsey.com/~/media/McKinsey/Industries/Advanced%20Electronics/Our%20Insights/How%20artificial%20intelligence%20can%20deliver%20real%20value%20to%20companies/MGI-Artificial-Intelligence-Discussion-paper.ashx)
+ [Notes from the AI Frontier: Insights from Hundreds of Use Cases](https://www.mckinsey.com/~/media/McKinsey/Featured%20Insights/Artificial%20Intelligence/Notes%20from%20the%20AI%20frontier%20Applications%20and%20value%20of%20deep%20learning/Notes-from-the-AI-frontier-Insights-from-hundreds-of-use-cases-Discussion-paper.ashx)
+ [Notes from the AI Frontier: Modeling the Impact on the World Economy](https://www.mckinsey.com/~/media/McKinsey/Featured%20Insights/Artificial%20Intelligence/Notes%20from%20the%20frontier%20Modeling%20the%20impact%20of%20AI%20on%20the%20world%20economy/MGI-Notes-from-the-AI-frontier-Modeling-the-impact-of-AI-on-the-world-economy-September-2018.ashx) 
+ [Notes from the AI Frontier:  Applying AI for Social Good](https://www.mckinsey.com/~/media/McKinsey/Featured%20Insights/Artificial%20Intelligence/Applying%20artificial%20intelligence%20for%20social%20good/MGI-Applying-AI-for-social-good-Discussion-paper-Dec-2018.ashx)

# How to Ask Questions

How to properly phrase a question is important in the workplace as well as in courses, such as ours, where much of the interaction is online.  Remember when formulating your question, *you get what you give*.  If you simply say, "I don't understand", then I have very little information to work with and so I can't help.  I'll just say to rewatch the relevant video.

Alternatively, if you say, "I calculated a stock's expected return using the CAPM and it is negative.  Can this be correct?".  Here you have (1) let me know what you have done, (2) your result, and (3) what you don't understand. In this case I can ask (1) which stock, and (2) is your beta negative, and proceed to explain whether your calculation is correct.

[Also, this is a helpful StackOverflow post.](https://stackoverflow.com/help/how-to-ask)

# Course Syllabus Disclosure Statement Fall 2021

> Accommodations for Students with Disabilities
> 
> In compliance with the Americans with Disabilities Act of 1990 and Section 504 of the Rehabilitation Act, SUNY Polytechnic Institute is committed to ensuring comprehensive educational access and accommodations for all registered students seeking access to meet course requirements and fully participate in programs and activities.  Students with documented disabilities or medical conditions are encouraged to request these services by registering with the Office of Disability Services.  Please request accommodations early in the semester, or as soon as you become registered with Disability Services, so that we have adequate time to arrange your approved academic accommodation/s.  Once Disability Services creates your accommodation plan, it is your responsibility to provide me a copy of the accommodation plan.
> 
> For information related to these services or to schedule an appointment, please contact the Office of Disability Services using the information provided below. 
> 
> Office of Disability Services 
> (315) 792-7170 
> Utica Campus 
> Peter J. Cayan Library, L145 
> Albany Campus 
> Suite 309, Students Services Office 
> NanoFab South 



